---
title: "DHS Pipeline (Quarto Orchestrator)"
author: "Tinashe M. Tapera"
format: gfm
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(targets)
library(tarchetypes)
library(here)
```

## Preamble

This project uses a notebook-driven pipeline architecture with three distinct layers:
function development, pipeline definition, and reporting.

Function-development notebooks live in a dedicated Quarto sub-project in `dev`. That
sub-project uses `sorting-hat` and `ripper` to extract selected R code chunks into
R module/package files. These generated functions define the reusable implementation
details of the pipeline.

The pipeline itself is defined in this notebook. Rather than being run as a normal
analysis report, this document is read by `tarchetypes::tar_tangle()`, which converts
top-level assignments in R chunks into `targets` objects. The `_targets.R` file then
loads this notebook as the literate source of the pipeline and `targets::tar_make()`
executes the resulting workflow.

Rendered reports, Quarto websites, and downstream analyses are kept separate from
pipeline definition and function extraction. They should consume completed pipeline
outputs from the `_targets/` store using `targets::tar_read()` or `targets::tar_load()`,
rather than redefining or rerunning the pipeline.

In short:

```text
function notebooks      -> generated R/package code
this target notebook    -> targets pipeline via tar_tangle()
targets pipeline        -> reproducible computed outputs in _targets/
reports/websites        -> downstream consumers of pipeline outputs
```

This separation lets us use notebook-driven development while avoiding accidental
coupling between code generation, pipeline execution, and rendered publication
content.

## `stagecoach` Input Tracking

The first thing to track is the stagecoach manifest file:

```{r}
#| format: file
#| packages: [here]

stagecoach_manifest <- here("stagecoach_manifest.yaml")
```

After running `stagecoach`, the input data were deposited:

```{r}
#| packages: [stringr, dplyr, here]

repo_files <- list.files(
    here(
      "data", 
      "inputs"
    ),
    recursive = TRUE
)
```

## Acknowledge File Types

As demonstrated in the `vignette("Linking and Tracking Raw Inputs")`,
we catalog the different types of DHS data files present in the raw data directory.

```{r}

raw_flat_dhs_files <- list_raw_flat_dhs(raw_data)
```


```{r}

raw_gps_dhs_files <- list_raw_gps_dhs(raw_data)
```


```{r}

raw_gps_covar_files <- list_raw_gps_covars(raw_data)
```


For each type above, we create a mapping to read the data in:
```{r}
#| packages:
#| - dplyr
#| - stringr
#| - purrr

raw_flat_dhs_files %>%
  str_subset(., "BR") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_BR

raw_flat_dhs_files %>%
  str_subset(., "HW") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_HW

raw_flat_dhs_files %>%
  str_subset(., "CR") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_CR

raw_flat_dhs_files %>%
  str_subset(., "HR") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_HR

raw_flat_dhs_files %>%
  str_subset(., "IR") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_IR

raw_flat_dhs_files %>%
  str_subset(., "KR") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_KR

raw_flat_dhs_files %>%
  str_subset(., "PR") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_PR

raw_flat_dhs_files %>%
  str_subset(., "WI") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_WI

raw_flat_dhs_files %>%
  str_subset(., "SQ") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_SQ

raw_flat_dhs_files %>%
  str_subset(., "MR") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_MR

raw_flat_dhs_files %>%
  str_subset(., "FW") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_FW
```


```{r}
#| packages: sf
#| pattern: map(raw_gps_dhs_files)
#| iteration: list

gps_data <- st_read(raw_gps_dhs_files)
```


```{r}
#| pattern: map(raw_gps_covar_files)
#| iteration: list

gps_covar_data <- load_gps_covars(raw_gps_covar_files)
```


## Climate Diarrhea Analysis

Olivia's analysis function to prepare the DHS data for climate-diarrhea work.
```{r}
#| packages:
#| - dplyr
#| - haven
#| - lubridate
#| - purrr
#| - tibble
#| - janitor
#| - glue
#| - rdhs

climate_diarrhea_data <- climate_diarrhea_analysis(dhs_data_HR)
```


To fetch this endpoint, after you've run `tar_make()`, you can use:

```r
tar_read(climate_diarrhea_data) #or
tar_load(climate_diarrhea_data)
```

