---
title: "DHS Pipeline (Quarto Orchestrator)"
author: "Tinashe M. Tapera"
format: gfm
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(targets)
library(tarchetypes)
library(here)

```

## Preamble

This project uses a notebook-driven pipeline architecture centered on `targets` and
Quarto. Rather than treating notebooks as polished reports or forcing exploratory
code into packages, notebooks are the primary development environment where the
computational workflow is discovered and recorded.

This notebook contributes to the project's `targets` pipeline. During project
execution, `tarchetypes::tar_tangle()` extracts the executable R code from pipeline
notebooks and converts top-level assignments into `targets` objects. The project's
`_targets.R` file assembles these targets into a directed acyclic graph (DAG), and
`targets::tar_make()` executes only the work that is out of date.

Not every target in the DAG represents a permanent part of the analysis. Throughout
development, exploratory, diagnostic, and intermediate targets are encouraged.
Target metadata (for example, descriptions) distinguish between development targets
that support exploration and the canonical targets that define the project's
reproducible workflow.

Rendered reports, Quarto websites, manuscripts, and downstream analyses are separate
consumers of the pipeline. They should read completed pipeline outputs from the
`_targets/` store using `targets::tar_read()` or `targets::tar_load()` rather than
recomputing the analysis themselves.

In short:

```text
exploratory notebooks   -> develop ideas and define targets
tar_tangle()            -> assemble the pipeline DAG
targets                 -> reproducible computational workflow
reports/websites        -> consume completed pipeline outputs
```

The guiding principle is simple: notebooks are where ideas are developed, while the
`targets` DAG is the durable record of the computations that matter. As an analysis
matures, important intermediate results become named targets, allowing the workflow
to emerge naturally from exploratory data analysis without requiring premature
software engineering.

## `stagecoach` Input Tracking

The first thing to track is the stagecoach manifest file:

```{r}
#| format: file
#| packages: [here]

stagecoach_manifest <- here("stagecoach_manifest.yml")
```

After running `stagecoach`, the input data were deposited:

```{r}
#| packages: [stringr, dplyr, here]

repo_files <- list.files(
    path = here(
      "data", 
      "inputs"
    ),
    recursive = TRUE,
    full.names = TRUE
)
```

## Acknowledge File Types

As demonstrated in the `vignette("Linking and Tracking Raw Inputs")`,
we catalog the different types of DHS data files present in the raw data directory.

```{r}
box::use(R/list_raw_flat_dhs[...])
raw_flat_dhs_files <- list_raw_flat_dhs(repo_files)
```


```{r}
raw_gps_dhs_files <- list_raw_gps_dhs(repo_files)
```


```{r}
raw_gps_covar_files <- list_raw_gps_covars(repo_files)
```


For each type above, we create a mapping to read the data in by year:
```{r}
#| packages:
#| - dplyr
#| - stringr
#| - purrr
library(dplyr)
library(tibble)
library(purrr)

dhs_flat_index <- tibble(path = raw_flat_dhs_files) %>%
  mutate(
    folder = dirname(path),
    extract = basename(folder),
    file = basename(path),
    recode = str_match(extract, "^MD([A-Z]{2})[0-9A-Z]{2}FL$")[, 2],
    survey_year = str_match(path, "/(?:DHS|MIS) ([0-9]{4}(?:-[0-9]{2})?)/")[, 2]
  ) %>%
  filter(!is.na(recode)) %>%
  distinct(survey_year, recode, extract, folder)

dhs_flat_index %>%
  mutate(data = map(folder, load_flat_dhs_data)) -> dhs_data_grp
```


```{r}
#| packages: sf
#| pattern: map(raw_gps_dhs_files)
#| iteration: list

gps_data <- st_read(raw_gps_dhs_files)
```


```{r}
#| pattern: map(raw_gps_covar_files)
#| iteration: list

gps_covar_data <- load_gps_covars(raw_gps_covar_files)
```


## Climate Diarrhea Analysis

Olivia's analysis function to prepare the DHS data for climate-diarrhea work.
```{r}
#| packages:
#| - dplyr
#| - haven
#| - lubridate
#| - purrr
#| - tibble
#| - janitor
#| - glue
#| - rdhs

#climate_diarrhea_data <- climate_diarrhea_analysis(dhs_data_HR)
```

## Exports

We export the data as a Data Package provided by Frictionless.

```{r}
dhs_data_grp %>%
    rename_all(~paste0("dhs_", .x)) %>%
    mutate(description = str_c(dhs_recode, "_", dhs_survey_year) %>% str_to_lower()) %>%
    group_by(description) %>%
    group_split() -> dhs_data_gp_spl

dp <- compose_data_package(dhs_data_gp_spl)
dhs_flat_dpkg <- write_package(dp, directory = here("data/outputs/dhs_flat_dpkg")); list.files(here("data/outputs/dhs_flat_dpkg"), recursive = TRUE, full.names = TRUE)
```

To fetch this endpoint, after you've run `tar_make()`, you can use:

```r
tar_read(climate_diarrhea_data) #or
tar_load(climate_diarrhea_data)
```
