---
title: "Exporting Curated Targets"
author: "Tinashe M. Tapera"
filters:
  - sorting-hat
  - ripper
extensions:
  sorting-hat:
    keep:
      - r
    verbose: true
  ripper:
    include-yaml: false
    output-dir: "../tests/testthat"
    output-name: "test-output_validation"
    script-links-position: "none"
format: gfm
---

In this notebook, we will export the curated targets from
our pipeline. This is an important step to ensure that 
the processed data is available for downstream analysis 
and can be easily consumed by other researchers. The export
will include some basic metadata about each dataset and
the date-time of export.

```{r}
library(here)
library(targets)
library(nanoparquet)
library(assertr)
library(frictionless)
library(sf)
library(stringr)

```

The DHS to GPS crosswalk is used for mapping DHS survey
points to the corresponding healthshed polygons. Meanwhile,
the climate and diarrhea data is used for analyzing the relationship
between climate variables and diarrhea incidence in Madagascar. It
was prompted by PR number [26](https://github.com/GoldenPlanetaryHealthLab/prospectorDHSMadagascar/pull/26).

```{r}
dhs_gps_2_healthshed_xwalk <- tar_read(dhs_gps_2_healthshed_xwalk, store = here("_targets")) %>%
  st_sf()
climate_diarrhea_data <- tar_read(climate_diarrhea_data, store = here("_targets"))
```

```{r}
dhs_gps_2_healthshed_xwalk
climate_diarrhea_data
```

We can create a data package for this pull request using `frictionless` 
to ensure that the data is well-documented and easily shareable.

```{r}
pr_datapkg <- frictionless::create_package()
```

We simply add the datasets to the package, and create the metadata with code:


```{r}
clm_d_schema <- frictionless::create_schema(climate_diarrhea_data)
clm_d_path <- here("data", "datalad", "8-geospatial-mapping", "climate_diarrhea_data", "climate_diarrhea_data.csv")
fs::dir_create(here("data", "datalad", "8-geospatial-mapping", "climate_diarrhea_data"))
clm_d_path %>%
  str_replace(".csv", ".parquet") %>%
  write_parquet(climate_diarrhea_data, .)
write.csv(climate_diarrhea_data, clm_d_path, row.names = FALSE)


xwalk_schema <- frictionless::create_schema(dhs_gps_2_healthshed_xwalk)
xwalk_path <- here("data", "datalad", "8-geospatial-mapping", "dhs_gps_2_healthshed_xwalk", "dhs_gps_2_healthshed_xwalk.csv")
fs::dir_create(here("data", "datalad", "8-geospatial-mapping", "dhs_gps_2_healthshed_xwalk"))
xwalk_path %>%
  str_replace(".csv", ".parquet") %>%
  write_parquet(dhs_gps_2_healthshed_xwalk, .)
write.csv(dhs_gps_2_healthshed_xwalk, xwalk_path, row.names = FALSE)

pr_datapkg <- pr_datapkg %>%
  add_resource(
    resource_name = "dhs_gps_2_healthshed_xwalk",
    data = xwalk_path,
    description = "Crosswalk between DHS GPS points and Madagascar healthshed polygons.",
    schema = xwalk_schema,
    pr_url = "https://github.com/GoldenPlanetaryHealthLab/prospectorDHSMadagascar/pull/26"
  ) %>%
  add_resource(
    resource_name = "climate_diarrhea_data",
    data = clm_d_path,
    description = "Climate and diarrhea data for analysis.",
    schema = clm_d_schema,
    pr_url = "https://github.com/GoldenPlanetaryHealthLab/prospectorDHSMadagascar/pull/26"
  )
```

Now we export the data package:

```{r}
pr_package_path <- here("data", "datalad", "8-geospatial-mapping")
pr_datapkg %>%
  write_package(directory = pr_package_path)
```

We currently have both forms (csv and parquet) for now; we will fix this in future.

Next, we push this dataset to our Datalad repository, which allows us to version control the data and make it accessible to other researchers.

