---
title: "Exporting Outputs"
format: gfm
filters:
  - coatless-quarto/sorting-hat
  - coatless-quarto/ripper
extensions:
  sorting-hat:
    keep:
      - r
    verbose: true
  ripper:
    include-yaml: false
    output-dir: "../R"
    output-name: "export"
    script-links-position: "none"
---

As the project matures, we export certain targets to disk to be used downstream:

1. A complete dataset gets packaged into a datapackage using `frictionless`
2. The data package is given a basic schema and metadata, and is exported to disk
3. Datalad is used to catalog the packages and push them to a remote registry

```{r}
#| sorting-hat: remove
library(targets)
library(frictionless)
library(here)
library(dplyr)
library(rdhs)
library(purrr)
library(tidyr)
library(haven)
library(stringr)
```

```{r}
#| sorting-hat: remove
tar_load(dhs_data_grp, store = here("_targets"))
```

```{r}
#| sorting-hat: remove
dhs_data_grp %>%
    rename_all(~paste0("dhs_", .x)) %>%
    mutate(description = str_c(dhs_recode, "_", dhs_survey_year) %>% str_to_lower()) %>%
    group_by(description) %>%
    group_split() -> dhs_data_gp_spl

dhs_wghts_spl <- dhs_weights %>%
    rename_all(~paste0("dhs_", .x)) %>%
    mutate(description = str_c(dhs_recode, "_", dhs_survey_year) %>% str_to_lower()) %>%
    group_by(description) %>%
    group_split()
```

```{r}
#| sorting-hat: keep
compose_data_package <- function(list_of_packages) {
    pkg <- create_package()
    
    list_of_packages %>%
        walk(
            .f = function(x) {
                if(ncol(x$dhs_data[[1]]) < 1) {
                    message("No data in ", unique(x$description), " skipping...")
                    return(NA)
                }
                
                dat <- data_and_labels(pluck(x$dhs_data, 1))$data %>%
                    zap_labels()
                
                labs <- data_and_labels(pluck(x$dhs_data, 1))$variable_names
                schema <- create_schema(dat)
                desc_lookup <- setNames(labs$description, labs$variable)

                schema$fields <- map(schema$fields, \(field) {
                    field$description <- unname(desc_lookup[field$name])
                    field
                })

                pkg <<- pkg %>%
                    add_resource(
                        data = dat,
                        resource_name = unique(x$description),
                        schema = schema
                    )
            }
        )

return(pkg)
}
```

```{r}
#| sorting-hat: remove
dp <- compose_data_package(dhs_data_gp_spl)
dp_wghts <- compose_data_package(dhs_wghts_spl)
```

```{r}
#| sorting-hat: remove
dhs_flat_dpkg <- {
    write_package(dp, directory = here("data/outputs/dhs_flat_dpkg")); list.files(here("data/outputs/dhs_flat_dpkg"), recursive = TRUE, full.names = TRUE)
}
dhs_wghts_dpkg <- {
    write_package(dp_wghts, directory = here("data/outputs/dhs_wghts_dpkg")); list.files(here("data/outputs/dhs_wghts_dpkg"), recursive = TRUE, full.names = TRUE)
}