---
title: "flat_build_targets_pipeline.Rmd empty"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
```

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```

## Preamble

The primary distribution mechanism of our data in the lab is via a
`targets` pipeline. `targets` is a powerful R package for
orchestrating data analysis pipelines. It helps manage dependencies,
ensures reproducibility, and optimizes performance by only rerunning
the parts of the pipeline that have changed.

In the lab, we separate the technical concerns of data storage, processing, analysis,
and reporting into distinct workspaces:

1. Storage: Each project's raw data is stored in a dedicated data repository on Google Drive (`Drive / <PROJECT> / 4. Datasets / <DATASET NAME>`).
  That dataset is _mirrored_ to FASRC via `rclone` for processing. To work on your
  own project, you must be on FASRC -- the pipeline will then _symlink_ the google drive data to your
  local "data/" directory using the `link_inputs()` function defined in the pipeline.
2. Processing and Analysis: The `targets` pipeline handles preliminary data processing
  and analysis. It reads raw data from the "data/" directory, processes it,
  and generates intermediate datasets and analysis results in your local workspace. To do this, simply:
    - Install this package in your local project space using `remotes::install_github()`.
    - Load the package using `library(your_package_name)`.
    - Run the `targets` pipeline by executing `build_targets_pipeline()` and then `tar_make()` in the R console. This will execute all defined targets, processing the data and generating clean datasets.
    - Access the processed data and analysis results stored in the `_targets/` directory using `tar_load()` or other `targets` functions.
3. Analysis: Once you've run the `targets` pipeline, you can access the processed data
  and analysis results stored in the `_targets/` directory. This data is then used
  for further analysis and reporting.
4. Reporting: Your modelling and analysis are conducted on your own in the project space,
  using the processed data outputs from the `targets` pipeline. You can create reports,
  visualizations, and other outputs based on the analysis results. I recommend using 
  Quarto for this purpose.
5. Sharing: Finally, when you are ready to share your work, use the respective project's
  `5. Data & Code Elements` Google Drive folder. This folder is intended for sharing
  intermediate data products — not raw or final data. You should always be able to reproduce
  any data product shared here by running the `targets` pipeline from raw data, and then rerunning
  your own analysis code.

# build_targets_pipeline

This is a convenience function that can build the targets pipeline for the user
in their local project. It does this by providing skeleton code for the targets file,
as well as the list of necessary targets in a separate Quarto file, which is converted
to targets objects via `tarchetypes::tar_tangle()`

```{r development-libraries}
library(glue)
library(readr)
```

```{r development-build_targets_pipeline}
# Prepare the code of your function here
```

```{r function-build_targets_pipeline}
#' build_targets_pipeline Title
#'
#' Dynamically build the targets pipeline for the project.
#' 
#' This function sets up the targets pipeline by defining a local `_targets.R` file
#' and using `tar_tangle()` to create the necessary targets from an Rmd file.
#' The Rmd file is, itself, this vignette — Pipeline Definition.
#' 
#' @importFrom glue glue
#' @importFrom here here
#' @return 1
#' @export
#'
#' @examples
build_targets_pipeline <- function(
  targets_notebook_name = "_target_notebook.Rmd", 
  targets_file_name = "_targets.R", 
  pkg_name = "DHSHarmonization", 
  vignette_name = "pipeline-definition.Rmd",
  overwrite = FALSE
  ) {
  
  # Define the path to the targets notebook definition and file
  targets_notebook <- here::here(targets_notebook_name)
  targets_file <- here::here(targets_file_name)

  # fetch this installed vignette
  # v <- vignette(vignette_name, package = pkg_name)
  # vignette_path <- fs::path(v$Dir, "doc", v$File)

  vignette_path <- system.file(vignette_name, package = pkg_name)

  # assert that the vignette exists
  if (!fs::file_exists(vignette_path)) {
    cli::cli_alert_danger("Vignette '{vignette_name}' not found in package '{pkg_name}'.")
    stop("Cannot proceed without the vignette file.")
  }

  # copy to local project
  fs::file_copy(vignette_path, targets_notebook, overwrite = overwrite)
  # remove irrelevant quarto directives
  lines <- readr::read_lines(targets_notebook) %>%
    stringr::str_subset("#\\| eval: no", negate = TRUE)
  writeLines(lines, targets_notebook)
  cli::cli_alert_success("Created local targets notebook file at: {targets_notebook}")

  # fetch the installed template for _targets.R
  template_path <- system.file("_targets.R", package = pkg_name)

  # copy to local project
  fs::file_copy(template_path, targets_file, overwrite = overwrite)
  cli::cli_alert_success("Created local targets file at: {targets_file}")

  # replace the tar_tangle() default to use the local notebook name
  if(targets_notebook_name != "_target_notebook.Rmd") {
    
    xfun::gsub_file(
      targets_file,
      pattern = "_target_notebook.Rmd",
      replacement = targets_notebook_name
    )
    cli::cli_alert_success("Updated targets file to use targets notebook: {targets_notebook_name}")
    
  }

  cli::cli_alert_info("To build the base pipeline, run `targets::tar_make()` in your R console. Use `targets::tar_read()` or `targets::tar_load()` to access target outputs.")
  cli::cli_alert_info("You can also edit the targets file to extend the pipeline as needed for your project.")

  return(invisible())
}
```

```{r examples-build_targets_pipeline}
#build_targets_pipeline()
```

```{r tests-build_targets_pipeline}
test_that("build_targets_pipeline works", {
  expect_true(inherits(build_targets_pipeline, "function"))
})
```

# Pipeline Definition

Below we define the actual targets of the pipeline using `tar_tangle()`.

## Config

A YAML config file is used to define parameters for the pipeline. At present,
the only important parameter is the path to the raw data repository on FASRC.

```{r, eval=FALSE}
#| format: file

cfg <- system.file("config.yml", package = "DHSHarmonization")
```

## Raw Data Linking

We use the function `link_inputs()` to symlink the raw data from the data repository
to your local "data/" directory. If successful, stores the full list of files in the raw data.


```{r, eval=FALSE}
#| format: file
#| error: stop
raw_data <- link_inputs(cfg_path = cfg)
```

## Acknowledge File Types

As demonstrated in the `vignette("Linking and Tracking Raw Inputs")`,
we catalog the different types of DHS data files present in the raw data directory.

```{r, eval=FALSE}
raw_flat_dhs_files <- list_raw_flat_dhs(raw_data)
```

```{r, eval=FALSE}
raw_gps_dhs_files <- list_raw_gps_dhs(raw_data)
```

```{r, eval=FALSE}
raw_gps_covar_files <- list_raw_gps_covars(raw_data)
```

For each type above, we create a mapping to read the data in:

```{r, eval=FALSE}
#| packages: [dplyr, stringr, purrr]
raw_flat_dhs_files %>%
  str_subset(., "BR") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_BR

raw_flat_dhs_files %>%
  str_subset(., "HW") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_HW

raw_flat_dhs_files %>%
  str_subset(., "CR") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_CR

raw_flat_dhs_files %>%
  str_subset(., "HR") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_HR

raw_flat_dhs_files %>%
  str_subset(., "IR") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_IR

raw_flat_dhs_files %>%
  str_subset(., "KR") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_KR

raw_flat_dhs_files %>%
  str_subset(., "PR") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_PR

raw_flat_dhs_files %>%
  str_subset(., "WI") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_WI

raw_flat_dhs_files %>%
  str_subset(., "SQ") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_SQ

raw_flat_dhs_files %>%
  str_subset(., "MR") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_MR

raw_flat_dhs_files %>%
  str_subset(., "FW") %>%
  map(~ load_flat_dhs_data(.x)) -> dhs_data_FW
```

```{r, eval=FALSE}
#| packages: [sf]
#| pattern: map(raw_gps_dhs_files)
#| iteration: list

gps_data <- st_read(raw_gps_dhs_files)
```

```{r, eval=FALSE}
#| pattern: map(raw_gps_covar_files)
#| iteration: list

gps_covar_data <- load_gps_covars(raw_gps_covar_files)
```

## Climate Diarrhea Analysis

Olivia's analysis function to prepare the DHS data for climate-diarrhea work.

```{r, eval=FALSE}
#| packages: [dplyr, haven, lubridate, purrr, tibble, janitor, glue, rdhs]

climate_diarrhea_data <- climate_diarrhea_analysis(dhs_data_HR)
```

To fetch this endpoint, after you've run `tar_make()`, you can use:

```r
tar_read(climate_diarrhea_data) #or
tar_load(climate_diarrhea_data)
```

```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_build-targets-pipeline.Rmd", vignette_name = "Pipeline Definition", check=FALSE, open=FALSE)
# Also copy the pipeline-definition.Rmd to inst/
fs::file_copy("vignettes/pipeline-definition.Rmd", "inst/pipeline-definition.Rmd", overwrite = TRUE)
```

