---
title: "flat_climate-diarrhea-analysis.Rmd empty"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
```

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```

# climate_diarrhea_analysis

This endpoint function develops the dataset for analyzing the relationship between climate variables
and diarrhea incidence, based on [Issue #2](https://github.com/Climate-Smart-Public-Health/DHSHarmonization/issues/2).

```{r}
library(targets)
library(here)
library(dplyr)
library(purrr)
library(rdhs)
library(janitor)
library(lubridate)
library(glue)
tar_load(dhs_data_HR, store = here("_targets/"))

```

We first load the DHS data from the household recode and filter the relevant variables from the request:

```{r}
dhs_data_HR %>% length()
```

There's 7 datasets; will they merge automatically?

```{r}
tryCatch({

  dhs_data_HR %>% rdhs::rbind_labelled()
}, error = function(e) {
  message("Error in merging datasets: ", e$message)
})
```

Nope, so we have to filter it manually:

```{r}
var_req_list <- c(
  "hv001",
  "hv008",
  "hv209",
  "hv013",
  "hv009",
  "hv201",
  "hv025",
  "hv205",
  "hv225",
  "hv237",
  "hv235",
  "hv238",
  "hv237b",
  "hv237a",
  "hv237z",
  "hv237f",
  "hv237x",
  "hv237e",
  "hv237c",
  "hv246f",
  "hv237d",
  "hv246b",
  "hv246d",
  "hv246c",
  "hv246",
  "hv246e",
  "hv232",
  "hv238a",
  "hv246f",
  "hv246g",
  "hv230a",
  "hv230b",
  "sh18a",
  "hv246a",
  "hv246f",
  "hv246b",
  "hv246g",
  "hv246h",
  "hv246d",
  "hv246c",
  "sh139c",
  "sh139b",
  "sh139a",
  "hv233",
  "hv232",
  "hv231",
  "hv246",
  "hv246h",
  "hv246i",
  "hv246j",
  "hv246k",
  "hv246a",
  "hv246f",
  "hv246a",
  "hv246i",
  "hv246h",
  "hv230",
  "hv246e",
  "sh22a",
  "sh138"
)
dhs_data_HR %>% 
  summarize_dhs_flat_dictionary() %>%
  mutate(description = janitor::make_clean_names(description)) -> dhs_dict_HR
```

```{r}
dhs_dict_HR %>%
  dplyr::filter(variable %in% var_req_list)
```

Now we need to filter each dataset for these variables and then merge them:

```{r}
dhs_data_HR %>%
  map(., function(df) {
    df %>%
      dplyr::select(hv000, hv006, hv007, hv008,dplyr::any_of(var_req_list))

  }) %>%
  map(., function(df) {
    data_and_labels(df) %>%
      pluck("dataset") %>%
      haven::as_factor()
  }) %>%
  bind_rows() %>%
  tibble() -> dhs_data_filtered_list
```

To rename the columns:

```{r}
rename_map <- setNames(dhs_dict_HR$variable, dhs_dict_HR$description)
dhs_data_filtered_list %>%
  rename(any_of(rename_map)) -> dhs_data_renamed
```

We should also manipulate the date of the interview:

```{r}
dhs_data_renamed %>%
  mutate(interview_date = my(glue("{month_of_interview} {year_of_interview}"))) %>%
  select(-c(month_of_interview, year_of_interview, date_of_interview_cmc)) -> dhs_data_final

skimr::skim(dhs_data_final)
```

Looks great! Let's put this in the function:

```{r function-climate_diarrhea_analysis}
#' climate_diarrhea_analysis
#'
#' Prepare and harmonize DHS household recode (HR) datasets for climate–diarrhea analysis.
#'
#' This function:
#' - Summarizes the flat DHS dictionary for the provided HR datasets and cleans variable descriptions.
#' - Builds a rename map from cleaned descriptions to original DHS variable names.
#' - Selects a set of required variables (always includes hv000, hv006, hv007, hv008) plus any variables supplied in
#'   `var_req_list` from each dataset in the input list.
#' - Converts labelled vectors to factors (via `data_and_labels()` and `haven::as_factor()`), binds rows across surveys,
#'   and returns a single tibble.
#' - Renames columns using the cleaned-description -> variable-name mapping, creates a parsed `interview_date`
#'   (from `month_of_interview` and `year_of_interview`), and drops intermediate interview date fields.
#'
#' @param dhs_data_HR A list of DHS household recode (HR) data frames / tibbles (flat format). Each element is a survey
#'   dataset. The function expects these to be in the flat DHS format and that a corresponding flat dictionary can be
#'   produced via `summarize_dhs_flat_dictionary()` and that `data_and_labels()` is available to extract labelled
#'   variables.
#' @param var_req_list Character vector of additional DHS variable names or cleaned descriptions to keep for the analysis.
#'   Defaults to a long list of common DHS household variables used in climate–diarrhea work (see function default).
#'
#' @return A tibble combining the selected variables from all input HR datasets. Columns are renamed according to the
#'   cleaned dictionary descriptions -> original variable names. A new column `interview_date` (Date) is created from
#'   `month_of_interview` and `year_of_interview`. Intermediate interview date columns
#'   (`month_of_interview`, `year_of_interview`, `date_of_interview_cmc`) are removed.
#'
#' @details
#' The function relies on the following workflow:
#' - `summarize_dhs_flat_dictionary()` is called on the input list to produce a dictionary; variable descriptions are
#'   cleaned with `janitor::make_clean_names()` to form keys for renaming.
#' - Each survey is subset to required variables using `dplyr::select()` and `dplyr::any_of()`.
#' - Labelled vectors are converted to R factors using `data_and_labels()` and `haven::as_factor()`.
#' - The per-survey datasets are row-bound together and returned as a tibble after renaming and date construction.
#'
#' @note
#' - The input list should contain DHS HR datasets in the expected flat format; missing expected fields may cause errors.
#' - This function uses functions from dplyr, purrr, janitor, haven, glue, and lubridate; ensure those packages are loaded.
#'
#' @examples
#' \dontrun{
#' # Assume `dhs_data_HR` is a named list of DHS HR data frames already loaded into the R session:
#' result <- climate_diarrhea_analysis(dhs_data_HR)
#'
#' # To request a custom set of variables:
#' vars <- c("hv001", "hv008", "hv201", "hv025")
#' result2 <- climate_diarrhea_analysis(dhs_data_HR, var_req_list = vars)
#' }
#'
#' @export
climate_diarrhea_analysis <- function(
  dhs_data_HR,
  var_req_list = c(
    "hv001",
    "hv008",
    "hv209",
    "hv013",
    "hv009",
    "hv201",
    "hv025",
    "hv205",
    "hv225",
    "hv237",
    "hv235",
    "hv238",
    "hv237b",
    "hv237a",
    "hv237z",
    "hv237f",
    "hv237x",
    "hv237e",
    "hv237c",
    "hv246f",
    "hv237d",
    "hv246b",
    "hv246d",
    "hv246c",
    "hv246",
    "hv246e",
    "hv232",
    "hv238a",
    "hv246f",
    "hv246g",
    "hv230a",
    "hv230b",
    "sh18a",
    "hv246a",
    "hv246f",
    "hv246b",
    "hv246g",
    "hv246h",
    "hv246d",
    "hv246c",
    "sh139c",
    "sh139b",
    "sh139a",
    "hv233",
    "hv232",
    "hv231",
    "hv246",
    "hv246h",
    "hv246i",
    "hv246j",
    "hv246k",
    "hv246a",
    "hv246f",
    "hv246a",
    "hv246i",
    "hv246h",
    "hv230",
    "hv246e",
    "sh22a",
    "sh138"
  )) {

  dhs_data_HR %>% 
    summarize_dhs_flat_dictionary() %>%
    mutate(description = janitor::make_clean_names(description)) -> dhs_dict_HR

  rename_map <- setNames(dhs_dict_HR$variable, dhs_dict_HR$description)
  
  dhs_data_HR %>%
    map(., function(df) {
      df %>%
        dplyr::select(hv000, hv006, hv007, hv008,dplyr::any_of(var_req_list))

    }) %>%
    map(., function(df) {
      data_and_labels(df) %>%
        pluck("dataset") %>%
        haven::as_factor()
    }) %>%
    bind_rows() %>%
    tibble() -> dhs_data_filtered_list

  dhs_data_filtered_list %>%
    rename(any_of(rename_map)) %>%
    mutate(interview_date = my(glue("{month_of_interview} {year_of_interview}"))) %>%
    select(-c(month_of_interview, year_of_interview, date_of_interview_cmc)) -> dhs_data_final

  return(dhs_data_final)
}
```

```{r examples-climate_diarrhea_analysis}
#climate_diarrhea_analysis()
```

```{r tests-climate_diarrhea_analysis}
test_that("climate_diarrhea_analysis works", {
  expect_true(inherits(climate_diarrhea_analysis, "function"))
})
```


```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_climate-diarrhea-analysis.Rmd", vignette_name = "Climate Diarrhea Analysis", check=FALSE, open=FALSE)
```

