---
title: "7-DHS-data-exploration-climate-and-general-foodborne-illness"
format: html
---

THis notebook addresses the addition of the following variables:

```{r}
library(readr)
library(here)
library(dplyr)
library(tibble)
library(rdhs)
library(purrr)
library(stringr)

vars <- here("data", "datalad", "7-DHS-data-exploration-climate-and-general-foodborne-illness", "DHS_explore_cjk.csv") %>%
   read_csv()
```

We pull in the main targets:

```{r}
library(targets)
tar_manifest(store = here("_targets"))
```

We'll pull the variables from each category in `vars`:

```{r}
unique(vars$category)
```

```{r}
hr <- tar_read(dhs_data_HR, store = here("_targets")) %>%
   enframe()
pr <- tar_read(dhs_data_PR, store = here("_targets")) %>%
   enframe()
wi <- tar_read(dhs_data_WI, store = here("_targets")) %>%
   enframe()
```

Each of these is a list of data frames, one for each year. We will want the
standard columns for each, and the select columns that have been requested in the `vars` file.

For HR, we will always select hv0**, and then the requested:

```{r}
hr_vars <- vars %>%
   filter(category == "HR") %>%
   pull(Variable) %>%
   c(paste0("hv", str_pad(0:100, 3, pad = "0")), .)
```

We'll do the same for the other categories:

```{r}
pr_vars <- vars %>%
   filter(category == "PR") %>%
   pull(Variable) %>%
   c(paste0("hv", str_pad(0:100, 3, pad = "0")), .)
wi_vars <- vars %>%
   filter(category == "WI") %>%
   pull(Variable) %>%
   c(paste0("hv", str_pad(0:100, 3, pad = "0")), .)
```

Now, we filter the dataframes:

```{r}
hr_filtered <- hr %>%
   mutate(foodborne_illness = map(value, ~ select(.x, any_of(hr_vars)))) %>%
   pull(foodborne_illness)

pr_filtered <- pr %>%
   mutate(foodborne_illness = map(value, ~ select(.x, any_of(pr_vars)))) %>%
   pull(foodborne_illness)

wi_filtered <- wi %>%
   mutate(foodborne_illness = map(value, ~ select(.x, any_of(wi_vars)))) %>%
   pull(foodborne_illness)
```

Now, we add a `frictionless` catalog view:

```{r}
library(frictionless)
```

```{r}
climate_foodborne_illness_hr <- reduce2(
  hr_filtered,
  seq_along(hr_filtered),
  function(pkg, x, i) {
    dat <- data_and_labels(x)$data %>%
         delabel_df()
    labs <- data_and_labels(x)$variable_names

    schema <- create_schema(dat)

    desc_lookup <- setNames(labs$description, labs$variable)

    schema$fields <- map(schema$fields, \(field) {
      field$description <- unname(desc_lookup[field$name])
      field
    })

    pkg %>%
      add_resource(
        data = dat,
        resource_name = paste0("hr_foodborne_illness_", i),
        schema = schema
      )
  },
  .init = create_package()
)

climate_foodborne_illness_pr <- reduce2(
  pr_filtered,
  seq_along(pr_filtered),
  function(pkg, x, i) {
    dat <- data_and_labels(x)$data %>%
      delabel_df()
    labs <- data_and_labels(x)$variable_names

    schema <- create_schema(dat)

    desc_lookup <- setNames(labs$description, labs$variable)

    schema$fields <- map(schema$fields, \(field) {
      field$description <- unname(desc_lookup[field$name])
      field
    })

    pkg %>%
      add_resource(
        data = dat,
        resource_name = paste0("pr_foodborne_illness_", i),
        schema = schema
      )
  },
  .init = create_package()
)

climate_foodborne_illness_wi <- reduce2(
  wi_filtered,
  seq_along(wi_filtered),
  function(pkg, x, i) {
    dat <- data_and_labels(x)$data %>%
      delabel_df()
    labs <- data_and_labels(x)$variable_names

    schema <- create_schema(dat)

    desc_lookup <- setNames(labs$description, labs$variable)

    schema$fields <- map(schema$fields, \(field) {
      field$description <- unname(desc_lookup[field$name])
      field
    })

    pkg %>%
      add_resource(
        data = dat,
        resource_name = paste0("wi_foodborne_illness_", i),
        schema = schema
      )
  },
  .init = create_package()
)
```

And we write those out to file


```{r}
write_package(climate_foodborne_illness_hr, here("data", "datalad", "7-DHS-data-exploration-climate-and-general-foodborne-illness", "climate_foodborne_illness_hr"))
write_package(climate_foodborne_illness_pr, here("data", "datalad", "7-DHS-data-exploration-climate-and-general-foodborne-illness", "climate_foodborne_illness_pr"))
write_package(climate_foodborne_illness_wi, here("data", "datalad", "7-DHS-data-exploration-climate-and-general-foodborne-illness", "climate_foodborne_illness_wi"))
```