---
title: "Geospatial Data Mapping"
format: html
---

This notebook extends the pipeline to incorporate
a geospatial mapping of data points. In line with issue
[#8](https://github.com/GoldenPlanetaryHealthLab/DHSHarmonization/issues/8),
we'll utilize `sf` to create a geospatial mapping
between the DHS data points and the points covered
in our healthshed. This will allow future researchers
to be able to run aggregation statistics of the DHS
to the healthshed level, which is more relevant for
public health decision-making.

```{r}
library(sf)
library(here)
library(dplyr)
library(ggplot2)
library(purrr)
library(targets)
library(testthat)

tar_load(gps_data, store = here("_targets"))
```

We'll start by taking a look at the existing GPS data:

```{r}
print(gps_data[[1]])
```

We have `r length(gps_data)` years of GPS data. Each survey
is tagged at a location with a DHSID identifier. What we want to do
is overlay these GPS points on the healthshed polygons, and assign
the point to the corresponding healthshed.

To do this, we'll first add a target to read in the healthshed shapefile:

```{r}
#| format: file
#| packages: [here]

mdg_healthshed_raw <- here("data", "inputs", "01_gold_mine", "mdg_healthsheds", "mdg_healthsheds2022.zip")
```

```{r}
#| packages: [sf]

mdg_healthsheds <- {
  
  tmp <- tempfile()
  unzip(mdg_healthshed_raw, exdir = tmp)
  shp_file <- list.files(tmp, pattern = "\\.shp$", full.names = TRUE, recursive = TRUE)
  st_read(shp_file)
  
}
```


Ensure both datasets use the same coordinate reference system (CRS):

```{r}

map(
  gps_data,
  ~ .x %>%
    st_crs()
)

```

Add this as a test:

```{r}
test_that("CRS of healthsheds and DHS GPS data match", {

  hlthshds <- tar_read(mdg_healthsheds, store = here("_targets"))
  gps <- tar_read(gps_data, store = here("_targets"), branches = 1)[[1]]

  expect_true(st_crs(hlthshds) == st_crs(gps))

})
```

Now we can perform a spatial join to map the GPS points to the healthshed polygons. We're using the `st_intersects` algorithm for 
this join:

```{r}
{
  joined <- st_join(gps_data[[1]], mdg_healthsheds, left = TRUE, join = st_intersects)

  print(joined %>% filter(!is.na(fs_uid)))
}
```

By doing this join, we retain `r nrow(joined)` rows of the original `r nrow(gps_data[[1]])` GPS points, which 
means that 3 points were not mapped to any healthshed. This could be due to the 
points being outside the boundaries of the healthsheds or due to inaccuracies 
in the GPS data. What if we use a different algorithm?

```{r}
{
  joined <- st_join(gps_data[[1]], mdg_healthsheds, left = TRUE, join = st_within)

  print(joined %>% filter(!is.na(fs_uid)))
}
```

No difference in using `within`. Following the illustration
available on [this medium post](https://mentin.medium.com/which-predicate-cb608b470471),
I'm going to go with my gut and assume we should be using `st_intersects` for 
this join, as it is more inclusive and allows for points that may be on the 
boundary of healthsheds to be included in the mapping.

Now that we know, what does it look like:

```{r}
map(
  gps_data,
  function(x) {
    year <- unique(x$DHSYEAR)
    pts <- x
    joined <- st_join(pts, mdg_healthsheds, left = FALSE)
    summary_counts <- joined %>% 
      st_drop_geometry() %>%
      group_by(fs_uid, fs_name) %>% 
      summarise(n_points = n(), .groups = "drop") %>%
      mutate(DHSYEAR = year)

    return(summary_counts)

  }
) %>%
  list_rbind() %>%
  left_join(mdg_healthsheds, ., by = c("fs_uid", "fs_name")) %>%
  ggplot() +
  geom_sf(aes(fill = n_points)) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90") +
  theme_minimal() +
  labs(fill = "# GPS points") +
  facet_wrap(DHSYEAR ~ ., nrow=2)
```

It seems that the healthsheds themselves are only sparsely
covered by the GPS points; which makes sense.

We'll provide this final table as a target output for future users:

```{r}
#| packages: [dplyr, sf, purrr]

gps_healthshed_mapping <- map(
  gps_data,
  function(x) {
    year <- unique(x$DHSYEAR)
    pts <- x
    joined <- st_join(pts, mdg_healthsheds, left = FALSE)
    joined
  }
) %>%
  list_rbind()
```

Now, the `gps_healthshed_mapping` target can be used 
in future analyses to aggregate DHS and climate data to the 
healthshed level, by referencing and joining your data
to this dataframe using the `fs_uid` column for healthshed
lookups, and to the DHSID column for DHS survey lookups.

```{r}
head(gps_healthshed_mapping)
```