---
title: "Loading All the DHS Data"
format: gfm
filters:
  - coatless-quarto/sorting-hat
  - coatless-quarto/ripper
extensions:
  sorting-hat:
    keep:
      - r
    verbose: true
  ripper:
    include-yaml: false
    output-dir: "../R"
    output-name: "load_data"
    script-links-position: "none"
---

```{r development-load}
#| sorting-hat: remove
# Load already included functions if relevant
#pkgload::load_all(export_all = FALSE)
library(here)
library(tidyverse)
library(purrr)
library(rdhs)
library(glue)
```

Here we demonstrate how we load and view the DHS datasets that will be
harmonized in this package. Each section builds the function needed
to read in the data, and then tests it to ensure it works as expected.
The functions are then implemented in the targets pipeline.

Some basic EDA is provided for you to help you understand what data
can be made available to you through a data request.

```{r}
#| sorting-hat: keep
library(targets)
library(here)
library(tidyverse)
library(purrr)
library(rdhs)
library(glue)
library(sf)
```

## load_flat_dhs_data

The goal of this section is a function that can read any
`flat` DHS dataset (e.g., BR, HR, IR, KR, PR, WI, CR, MR, FW, SQ) and return it as a tibble.

So long as the data has a flat structure, it should be straightforward to
load it into R and view it using the `rdhs` package. The structure is taken from
this [DHS page](https://dhsprogram.com/data/File-Types-and-Names.cfm#CP_JUMP_10334).

```
Example:
To give an example of how distribution files for a survey are organized, the following table shows the available files, along with the names that they are given for the Kenya 2003 DHS survey.
Kenya 2003 DHS Survey
 	ASCII File Types	Software-Specific Data File Types
Unit of Analysis	Hierarchical	Flat	SAS	SPSS	Stata
Households	 	KEHR42FL.ZIP	KEHR42SD.ZIP	KEHR42SV.ZIP	KEHR42DT.ZIP
Household Members	 	KEPR42FL.ZIP	KEPR42SD.ZIP	KEPR42SV.ZIP	KEPR42DT.ZIP
Women	KEIR42.ZIP	KEIR42FL.ZIP	KEIR42SD.ZIP	KEIR42SV.ZIP	KEIR42DT.ZIP
Men	KEMR42.ZIP	KEMR42FL.ZIP	KEMR42SD.ZIP	KEMR42SV.ZIP	KEMR42DT.ZIP
Births	 	KEBR42FL.ZIP	KEBR42SD.ZIP	KEBR42SV.ZIP	KEBR42DT.ZIP
Children	 	KEKR42FL.ZIP	KEKR42SD.ZIP	KEKR42SV.ZIP	KEKR42DT.ZIP
Couples	 	KECR42FL.ZIP	KECR42SD.ZIP	KECR42SV.ZIP	KECR42DT.ZIP
HIV Test Results	KEAR42.ZIP	KEAR42FL.ZIP	KEAR42SD.ZIP	KEAR42SV.ZIP	KEAR42DT.ZIP

The following reference tables contain the descriptions for the four different types of filename codes (country, data type, data version, and file format).

```

In our case, we have the following:

```
tree -L 3 data/DHS\ Data/
```

```
data/DHS Data/
├── DHS 1992
│   ├── MDBR21FL
│   │   ├── MDBR21FL.DAT
│   │   ├── MDBR21FL.DCF
│   │   ├── MDBR21FL.DCT
│   │   ├── MDBR21FL.DO
│   │   ├── MDBR21FL.frq
│   │   ├── MDBR21FL.frw
│   │   ├── MDBR21FL.MAP
│   │   ├── MDBR21FL.SAS
│   │   └── MDBR21FL.SPS
│   ├── MDHR21FL
│   │   ├── MDHR21FL.DAT
│   │   ├── MDHR21FL.DCT
│   │   ├── MDHR21FL.DO
│   │   ├── MDHR21FL.FRQ
│   │   ├── MDHR21FL.FRW
│   │   ├── MDHR21FL.MAP
│   │   ├── MDHR21FL.SAS
│   │   └── MDHR21FL.SPS
│   ├── MDHW21FL
│   │   ├── MDHW21FL.DAT
│   │   ├── MDHW21FL.DCF
│   │   ├── MDHW21FL.DCT
│   │   ├── MDHW21FL.DO
│   │   ├── MDHW21FL.MAP
│   │   ├── MDHW21FL.SAS
│   │   ├── MDHW21FL.SPS
│   │   └── MERGE.DOC
│   ├── MDIR21FL
│   │   ├── MDIR21FL.DAT
│   │   ├── MDIR21FL.DCT
│   │   ├── MDIR21FL.DO
│   │   ├── MDIR21FL.DOC
│   │   ├── MDIR21FL.FRQ
│   │   ├── MDIR21FL.FRW
│   │   ├── MDIR21FL.MAP
│   │   ├── MDIR21FL.SAS
│   │   └── MDIR21FL.SPS
│   ├── MDKR21FL
│   │   ├── MDKR21FL.DAT
│   │   ├── MDKR21FL.DCT
│   │   ├── MDKR21FL.DO
│   │   ├── MDKR21FL.DOC
│   │   ├── MDKR21FL.FRQ
│   │   ├── MDKR21FL.MAP
│   │   ├── MDKR21FL.SAS
│   │   └── MDKR21FL.SPS
│   ├── MDPR21FL
│   │   ├── MDPR21FL.DAT
│   │   ├── MDPR21FL.DCT
│   │   ├── MDPR21FL.DO
│   │   ├── MDPR21FL.FRQ
│   │   ├── MDPR21FL.FRW
│   │   ├── MDPR21FL.MAP
│   │   ├── MDPR21FL.SAS
│   │   └── MDPR21FL.SPS
│   └── MDSQ21FL
│       ├── MDSQ21FL.DAT
│       ├── MDSQ21FL.DCT
│       ├── MDSQ21FL.DO
│       ├── MDSQ21FL.MAP
│       ├── MDSQ21FL.SAS
│       └── MDSQ21FL.SPS
```

So, for the generic ones, we know that we have BR (Birth rates), HR (Household), HW (height and weight), IR (individual),
KR (kids), PR (household member), and SQ (Service Availability). We also have WI (wealth), CR (couples), MR (men's), 
GE (Geographic Data), GC (geospatial covariates), FW (Fieldworker), and MIS data (malaria response survey).

To read in the generic flat files, we can use the `rdhs` package. This has to be done by first zipping
the files to a temporary location, and then using the `rdhs::read_flat()` function.

```{r}
#| sorting-hat: remove
library(rdhs)
library(dplyr)
library(here)
library(zip)

ex_data <- here("data", "DHS Data", "DHS 1992", "MDIR21FL")
temp_zip <- tempfile(fileext = ".zip")

zip::zipr(zipfile = temp_zip, files = list.files(ex_data, full.names = TRUE))
file.exists(temp_zip)
```

```{r}
#| sorting-hat: remove
dhs_data <- rdhs:::read_dhs_flat(temp_zip)
tibble(dhs_data)
```

We can see that the variables do come with labels and attributes. This is
useful information for understanding the data.

```{r}
#| sorting-hat: remove
get_variable_labels(dhs_data) %>%
  head()
```

So, to read in a generic flat file from DHS, we can define the function `load_flat_dhs_data()`:

```{r function-load_flat_dhs_data}
#| sorting-hat: keep
#' Load a flat DHS dataset from a folder
#'
#' This function zips the contents of the provided folder and uses rdhs:::
#' read_dhs_flat to read a DHS flat file archive into R. It is intended for
#' DHS flat file folders (e.g. MDIR21FL) that contain the standard DHS flat
#' data files.
#'
#' @param folder_path Character. Path to a folder containing DHS flat files to
#'   be zipped and read.
#' @return A data.frame/tibble containing the DHS flat data as returned by
#'   rdhs:::read_dhs_flat.
#' @export
#' @examples
#' ex_data <- here::here("data", "DHS Data", "DHS 1992", "MDIR21FL")
#' df <- load_flat_dhs_data(ex_data)
#' head(df)
load_flat_dhs_data <- function(folder_path) {

  temp_zip <- tempfile(fileext = ".zip")
  zip::zipr(zipfile = temp_zip, files = list.files(folder_path, full.names = TRUE))
  dhs_data <- rdhs:::read_dhs_flat(temp_zip)
  return(dhs_data)

}
```

```{r examples-load_flat_dhs_data}
#| sorting-hat: remove
ex_data <- here("data", "DHS Data", "DHS 1992", "MDIR21FL")
load_flat_dhs_data(ex_data) %>%
  tibble() %>%
  head()
```

We can also test that this works for the malaria response survey data:

```{r}
#| sorting-hat: remove
ex_data <- here("data", "DHS Data", "MIS 2016", "MDIR71FL")

load_flat_dhs_data(ex_data) %>%
  tibble() %>%
  head()
```

Because the different kinds of surveys have different variable
structures, we will read each type separately.

To do this, we'll read in a list of file paths,
and then map over them to read them in.

In the targets pipeline, we've assigned the BR files to
a target called `dhs_data_BR`, which has multiple files per target.

```{r}
#| eval: false
#| sorting-hat: remove
tar_read(dhs_data_BR, store = here("_targets"))[[1]] -> ex_data_1
tar_read(dhs_data_BR, store = here("_targets"))[[5]] -> ex_data_2

tibble(ex_data_1)
```

```{r}
#| eval: false
#| sorting-hat: remove
tibble(ex_data_2)
```

So even though they come from the same type of survey,
the variables are different and cannot be easily merged with
the `rhds::rbind_labelled()` function, because they change year over year:

```{r, error=TRUE}
#| sorting-hat: remove
# try to merge all of the BR datasets
all_br_data <- tar_read(dhs_data_BR, store = here("_targets"))
tryCatch({
  rdhs::rbind_labelled(all_br_data) %>% tibble()
}, error = function(e) {
  message("Error merging BR datasets: ", e$message)
  NULL
})
```

So with that in mind, we should take a look at the actual
variable dictionaries for each survey type to plan how we will
eventually harmonize and release them (descriptions are provided
by ChatGPT and verified with DHS documentation available [here](https://preview.dhsprogram.com/pubs/pdf/DHSG1/Guide_to_DHS_Statistics_DHS-7_v2.pdf)).

## summarize_dhs_flat_dictionary

The function `summarize_dhs_flat_dictionary()` will read in all of the flat files of a given survey type (e.g., BR, HR, IR)
and summarize the variable names.

  
```{r function-summarize_dhs_flat_dictionary}
#| sorting-hat: keep
#' Summarize DHS flat-file variable dictionaries
#'
#' Summarize variable names and descriptions across a list of DHS survey datasets.
#'
#' @param all_survey_data A list of survey datasets to summarize.
#'
#' @return A tibble of variable descriptions with counts across surveys.
#'
#' @export
summarize_dhs_flat_dictionary <- function(all_survey_data){
  
  n_surveys <- length(all_survey_data) # shows the number of branches   

  map(
    all_survey_data,
    ~ get_variable_labels(.x) %>%
      as_tibble()
    ) %>%
    list_rbind() %>%
    group_by(description, variable) %>%
    summarise(
      description = first(description),
      n = n()
    ) %>%
    arrange(-n) -> var_descriptions

  return(var_descriptions)
}
```
  
  
