---
title: "DHS Pipeline"
author: "Tinashe M. Tapera"
filters:
  - sorting-hat
  - ripper
extensions:
  sorting-hat:
    keep:
      - r
    verbose: true
  ripper:
    include-yaml: false
    output-dir: "../tests/testthat"
    output-name: "test-pipeline_orchestrator"
    script-links-position: "none"
format: gfm
---

Here we document, create, and run the pipeline all in
one file thanks to Quarto and `targets`.

```mermaid

{{< include "pipeline.mmd" >}}

```


For the backend, we'll use this Quarto notebook as the 
"orchestrator," meaning that the pipeline runs any time
this notebook is executed/rendered. First, let's get 
the script files from the inputs directory:

```{r}
devtools::load_all()
```

```{r}
library(stringr)
library(testthat)
library(dplyr)
library(yaml)
library(purrr)
library(here)
library(targets)
```

The inputs were provided by stagecoach. The manifest for
the stagecoach run is in `stagecoach_manifest.yml` if you want to check it out.

```{r}
#| format: file
#| packages: [here, dplyr, yaml, purrr]

manifest <- here("stagecoach_manifest.yml")
```

```{r}
#| packages: [stringr, dplyr, here]

repo_files <- list.files(
    here(
      "data", 
      "inputs"
    ),
    recursive = TRUE
)
```

